---
title: "Lab 10 - Multilevel Models"
output: 
  tufte::tufte_html:
    css: 
    tufte_variant: "envisioned"
    highlight: pygments
    fig_height: 10
    toc: true
    toc_depth: 1
execute: 
  message: false
  warning: false
format: 
  html:
    code-fold: true
    code-overflow: wrap
---

# Lab 10: MLM

Today's lab uses data from a study conducted by Coyne et al. (2019), which examined the impact of a high-quality, evidence-based vocabulary instruction intervention in kindergarten. The data consists of 1,428 students who were nested in 233 clusters or classrooms (`clu_id`).

For the at-risk youth in the classroom (*N* = 6), half were allocated to treatment and the other half to control. The treatment group received supplemental small-group vocabulary instruction in addition to the whole-group instruction, while the control group received only whole-group vocabulary instruction. Since the observations were not independent (due to students being nested within classrooms), the researchers needed to account for this in their analysis.

The main question that the researchers aimed to answer was whether the supplemental small-group kindergarten vocabulary instruction intervention increased students' knowledge of the vocabulary words taught in the intervention. To measure vocabulary knowledge, the researchers used an ETW (ETW_SpringK) assessment, which evaluated students' ability to explain the meaning of a given word. The assessment was administered after the intervention concluded, in the spring of kindergarten. In the sample, ETW scores ranged from 0 to 52 (the maximum score), with a mean of 13.65 and a standard deviation of 11.10. To answer the research question, the researchers used two fixed effects and their interaction: `TRT` (1 = treatment and 0 = control) and PPVT (Peabody Picture Vocabulary Test, which measures students' vocabulary before the intervention; `PPVT_FallK`).

> Coyne, M. D., McCoach, D. B., Ware, S., Austin, C. R., Loftus-Rattan, S. M., & Baker, D. L. (2019). Racing Against the Vocabulary Gap: Matthew Effects in Early Vocabulary Instruction and Intervention. *Exceptional Children*, *85*(2), 163--179. <https://doi.org/10.1177/0014402918789162>

## Load packages

```{r}
library(tidyverse)
library(lme4) # lmer mixed modeling
library(emmeans) # testing marginal means 
library(ggeffects) # plotting 
library(interactions) # plotting interactions and testing 

```

## Load data

```{r}
# read in data file 'Ch3_MLM_R_reduced.csv`

```

## Lab 

1.  Contrast code the treatment variable

```{r}

```

2.  Center **`PPVT_Fallk.`** While we did not talk about centering. There are two kinds of centering you can do in MLMs. One is cluster-mean centering ($x_{ij}−x_j$) where we subtract each person's cluster mean from each observation in that cluster. There is also grand-mean centering ($x_i−x$)subtract the overall mean from each observation). Group mean centering level-1 predictors allows us to test the pure within-cluster effect without confounding from between-cluster effects. Use the code from here: https://psy504-advanced-stats.netlify.app/slides/mlm/mlm-speaker#/group-mean-center-in-r to group mean center. For more information on centering I am including a link to a blog post by Patrick Curran: https://centerstat.org/centering/.

```{r}

```

3.  How much missing data is in our dataset? What variables have missing data and how much?

    ::: callout-note
    If the missing data is only on the DV, multilevel modeling is generally pretty good at handling it. Missingness on the predictors is another story...
    :::

```{r}

```

4.  Create two plots: One for the relationship between PPTV and EWR and one for the relationship between TRT and EWR. Make sure you plot the between -cluster (class) variability when you graph these.

    ```{r}


    ```

5.  Fit an unconditional means (null model)

```{r}

```

-   What is the ICC? Calculate it by hand from the output.

```{r}

```

-   Is multilevel modeling warranted here?

5.  Build up from the last model. Fit a model that includes all main effects (level 1 variables)

```{r}


```

6.  Fit a model that includes the fixed interaction between the level-1 variables.

```{r}


```

7.  Fit the maximal model warranted by this design

```{r}

```

8.  Compare the models. Which one is the best?

```{r}



```

9.  With the maximal model, test if the interaction is warranted in the model.

```{r}

```

10. Using the maximal model:

-   Test model assumptions

```{r}


```

-   Which fixed effects are significant? Follow these up with the appropriate tests (e.g., simple slopes analysis if interaction is significant).

```{r}



```

```{r}

```

-   Evaluate the variance components of your model.

    ```{r}

    ```

    -   Calculate the pseudo-$R^2$ to explain how much of the variance the fixed effects portion and random parts explain

        ```{r}


        ```

-   Visualize the random intercepts and slopes in your model

    ```{r}


    ```

-   Create a table with a summary of your model information

```{r}

```

11. Visualize the interaction between treatment and PPVT score on ETW (I would check out the `interactions` package. They have some nice features for plotting interactions.)

    ```{r}
    ```

12. Write-up the results of your MLM model in a paragraph or two
